---
output:
  pdf_document: default
  html_document: default
---

```{r}

data <- read.csv("fish-market-data.csv", fileEncoding="UTF-8-BOM")

```

# Unique Data Point Introduction 

A datapoint can be introduced which is an outlier and has high leverage. The reason is that these data points often exist in real world data due to error or incorrect collection, influence outside of the model, or it may be legitimate. It is up to the statistician to make a judgement call and decide on whether the data point should be kept or discarded.

 ```{r}

data <- rbind(data.frame(Species = 'Perch', Weight = 2000, Length1= 40.9, Length2 = 41.2, Length3 = 40.8, Height = 13.11, Width = 8.2), data)

```

# Data Cleaning

Manually browsing the data, it seems pretty clean. Some issues came up in the descriptive analysis though. In the summary data, one or more observations have zero weight.
 
 ```{r}
 
 data[data$Weight == 0, ]
 
  ```
  
  Clearly, a fish cannot have zero Wight. Since its other attributes seem fine, I don't think removing the observation is necessary. A better option would be to interpolate for Wight given that the Species type is a Roach. Weight has a similar relationship to all other variables, so ordering by one of these and using it as a basis for linear interpolation should be fine. Even though we see currently see that the relationship is not quite linear, interpolating in this way should be fine and not significantly impact the model results.
  
 ```{r}
 
roach_data <- data[data$Species == 'Roach', ]
rownames(roach_data) <- 1:nrow(roach_data)
zero_index <- as.numeric(rownames(roach_data[roach_data$Weight == 0, ]))
pred_val <- roach_data[c(zero_index), 3]

roach_data <- roach_data[-c(zero_index),]
roach_lm <- lm(Weight ~ Length1, data = roach_data)

inter_value <- predict(roach_lm, data.frame(Length1 = pred_val))
data[data$Weight == 0, 2] <- inter_value

  ```
  
From the descriptive analysis it was also evident that the data is in different units. To make the regression analysis simpler a unit normal scaling can be performed.

 ```{r}
 
scaled_data <- as.data.frame(scale(data[,-1], center = TRUE, scale = TRUE))

```

This is also a good time to setup our indicator variable species.

 ```{r}
 
indicator_data <- fastDummies::dummy_cols(data['Species'], select_columns = "Species")
indicator_data$Species <- NULL
indicator_data$Species_Bream <- NULL

```

Combine the scaled and indicator data.

 ```{r}
 
data <- data.frame(indicator_data, scaled_data)

```

Output the new dataframe as a .csv so it can be used in the regression analysis. 

 ```{r}
 
write.csv(data, 'cleaned_data_2.csv', row.names = FALSE)
write.csv(scaled_data, 'cleaned_data_scaled_only_2.csv', row.names = FALSE)

```



